---
title: "Non-linear Least Squares"
author: "Natalie Smith"
format: 
  html:
    code-fold: true
    toc: true
    number-sections: true
    embed-resources: true
theme: Yeti
editor: visual
execute:
  echo: true
  message: false
  warning: false
---

# Overview:

Farmers need to understand the biology of plants and their responses to fertilizers to maximize yield. This analysis aims to help farmers predict yields by running non-linear least squares on experimental growth data for three grains in Greece. Additionally, it assesses the response of the grains to fertilizer inputs.

Data:

The dataset contains experimental growth data with the following variables: [-]{.underline} DOY: Day of Year [-]{.underline} Block: Location [-]{.underline} Input: irrigation & Fertilizer application (High or Low) [-]{.underline} Crop: Type of crop (Fiber sorghum, Sweet sorghum, or Maize) [-]{.underline} Yield: Biomass

```{r}
library(tidyverse)
library(knitr)
library(broom)
library(kableExtra)
library(nlraa)
library(janitor)
library(patchwork)
```

```{r}
# Access the data from SM package in NLRAA
data <- sm
#clean the data
crop_df <- data %>% 
  janitor::clean_names()
```

<!-- -->

-   Crop:

    -   \(F\) fiber sorghum

    -   (S )sweet sorghum

    -   \(M\) maize

-   Yield: Biomass

Briefly describe what each parameter generally means:

-   Y is the response variable (e.g., biomass)

    -   Ymax is the asymptotic or the maximum Y value

-   t is the explanatory variable (e.g., day of year)

-   tm is the inflection point at which the growth rate is maximized

-   te is the time when Y = Yasym

Notes from paper: \* Three crops \* two levels of agronomic input \* four blocks \* = 24 possible combinations

### Create a Model:

Write a function in R to model the Beta Function. 

Create a plot similar to Fig. 4 (biomass over time) and use the definitions of the different parameters to find some potential starting parameter values for your nls analysis.

In your final report include the code, but NOT the plot you used to feed guesses.

```{r}
# Plot
guess_plot <- ggplot(crop_df, aes(x = doy, y = yield, color = "data")) +
  geom_point() +
  geom_smooth()+
  facet_grid(. ~ input, scales = "free_y", space = "free_x")+
  scale_color_manual(values = c("data" = "black", "model" = "#FF5733")) +
  labs(y = "Biomass (mg/ha)", x = "Day of Year", color = " ")+
  theme_minimal()

guess_plot
```

```{r}

# Create beta function:
beta<- function(ymax, te, tm, t){
  y <- ymax * (1 + (te - t) / (te - tm)) * (t / te)^(te / (te - tm))
  return(y)
}

# Guesses:
ymax_guess <- 20
te_guess <- 240
tm_guess <- 200

# Add inital guesses:
beta_test <- crop_df %>% 
  mutate(sim=beta(ymax = 20, te = 240, tm = 220, t= doy))


```

Model Fitting and Evaluation:

(one crop)

```{r}

# Filter to keep observations from the sorghum fields with high inputs
crop_filter <- crop_df %>% 
  filter((crop == "S") & input == 2)

```

Run NLS

```{r}
nls_1 <- nls(
  formula = yield ~ beta(ymax, te, tm, doy),
  data = crop_filter,
  start = list(ymax = ymax_guess, te = te_guess, tm = tm_guess),
  trace = TRUE)
```

Cute table:

```{r}
# Create a summary table 
summary_table <- coefficients(summary(nls_1))

# Basic table
kable(summary_table)

# Fancy table
kable(summary_table, format = "html", align = "lccc", digits = 3,
      col.names = c("Parameter", "Estimate", "Standard Error", "T-value", "P-value"))

```

How well the model fits the data:

```{r}
# Model visualization
crop_predict <- crop_filter %>% 
  mutate(predict=predict(nls_1,newdata=.))

ggplot(data=crop_predict)+
  geom_point(aes(x=doy, y=yield))+
  geom_path(aes(x=doy, y=predict),color="#FF5733")+
  labs(y = "Biomass (mg/ha)", x = "Day of Year", color = " ")+
  theme_minimal()
  
```

tbh, looks good?

Replicate the analysis using Purrr:

Did this do anything?

```{r}
#Define a new function to pass along the nls calls

all_nls <- function(test_df) {
  nls(yield ~ beta(ymax, te, tm, doy),
      data = test_df,
      start = list(ymax = ymax_guess, te = te_guess, tm = tm_guess))
}

# Replicate w Purrr

all_crops <- crop_df %>%
  group_by(block,input,crop) %>% 
  nest() %>% 
  mutate(nls_model = map(data, ~ all_nls(.x))) %>% 
  mutate(predictions = map2(nls_model, data, ~ predict(.x, newdata = .y))) %>% 
  mutate(rmse = map2_dbl(predictions, data, ~ Metrics::rmse(.x, .y$yield))) %>% 
  mutate(smooth = map(nls_model, ~predict(.x, newdata = list(doy = seq(147,306)))))

all_crops


```

Create a table

```{r}

rmse_table <- all_crops %>% 
  group_by(crop) %>% 
  summarize(rmse = min(rmse))

low_rmse <- all_crops %>% 
  filter(rmse %in% rmse_table$rmse)

low_rmse_M <- broom::tidy(low_rmse$nls_model[[1]]) %>% mutate(crop="M")
low_rmse_S <- broom::tidy(low_rmse$nls_model[[2]]) %>% mutate(crop="S")
low_rmse_F <- broom::tidy(low_rmse$nls_model[[3]]) %>% mutate(crop="F")

low_rmse_combined <- bind_rows(low_rmse_M,low_rmse_S,low_rmse_F)

kable(low_rmse_combined)
  
```

```{r}
# Unnest predictions from data and clean maize data
un_df <- all_crops %>% 
  filter(block==1) %>% 
  tidyr::unnest(smooth) %>% 
  mutate(doy=seq(147,306)) %>% 
  filter(!(doy>263 & crop=="M"))

# Create a dataframe to add corn data
hi_filter <- crop_df %>% 
  filter(block == 1 & input == 2)

low_filter <- crop_df %>% 
  filter(block == 1 & input == 1)

```

```{r}
#Make dumb graphs

hi_plot <- un_df %>%
  filter(block == 1 & input == 2) %>%
  ggplot() +
  geom_point(data = hi_filter, aes(x = doy, y = yield, shape = crop)) +
  geom_line(aes(x = doy, y = smooth, linetype = crop)) +
  labs(y = " ", x = " ", color = " ")+
  theme_minimal()

hi_plot


low_plot<-un_df |> 
  filter(block==1 & input==1) |> 
  ggplot()+
  geom_point(data=low_filter,aes(x=doy,y=yield,shape=crop))+
  geom_line(aes(x=doy,y=smooth,linetype=crop))+
  labs(y = " ", x = " ", color = " ")+
  theme_minimal()

low_plot


```

Combine Plots:

```{r}
library(patchwork)

plot1 <- hi_plot + 
  theme(axis.text.y = element_text(size = 8),
        axis.text.x = element_text(size = 8),  
        axis.title.x = element_blank(),
        axis.title.y = element_blank()) 


plot2 <- low_plot +
   theme(axis.text.y = element_text(size = 8),
        axis.text.x = element_text(size = 8),  
        axis.title.x = element_blank(),
        axis.title.y = element_blank()) 

combine_plots <- plot1 +plot2 + plot_layout(guides = "collect")

combine_plots


```
